---
title: "2019 Index: Drug overdoses"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = TRUE,
	message = FALSE,
	warning = FALSE,
	fig.showtext = TRUE
)
```

```{r}
library(tidyverse)
library(camiller)
library(cwi)
library(showtext)
library(sf)
library(tidycensus)
library(lubridate)
library(tsibble)
library(feasts)
source("../_utils/misc_functions.R")
```

Copied from health_equity repo (copied from 2019index)
```{r}
cluster_lst <- ct5_clusters %>%
  split(.$cluster) %>%
  map(pull, town)
```

```{r eval=FALSE}
pop_age <- multi_geo_acs("B01001", regions = cluster_lst) %>%
  label_acs() %>%
  separate(label, into = c("total", "sex", "age"), sep = "!!") %>%
  filter(!is.na(age)) %>%
  extract(age, into = c("min_age", "max_age"), regex = "^(\\d+)?[\\sA-Za-z]+(\\d+)?") %>%
  mutate_at(vars(ends_with("_age")), as.numeric) %>%
  replace_na(list(min_age = 0)) %>%
  mutate(age_brk = cut(min_age, breaks = c(0, seq(10, 85, by = 5), 100), include.lowest = TRUE, right = FALSE) %>%
           fct_relabel(age_brks)) %>%
  group_by(level, name = NAME, age_brk) %>%
  summarise(pop = sum(estimate))

saveRDS(pop_age, "../raw_data/pop_by_age.rds")

pop_read <- c(total = "", white = "H", black = "B", latino = "I") %>%
  map_chr(~paste0("B01001", .)) %>%
  map(~multi_geo_acs(table = ., year = 2018, regions = cwi::regions[main_regions])) %>%
  map(filter, !str_detect(level, "counties")) %>%
  map(label_acs)
saveRDS(pop_read, "../raw_data/pop_by_age_race_sex.rds")
```

```{r}
pop_age <- readRDS("../raw_data/pop_by_age.rds") %>%
  ungroup() %>%
  mutate(level = fct_relabel(level, str_remove, "^\\d+_"))

drugs_indiv <- read_csv("../output_data/overdoses_indiv_2012_2020.csv") %>%
  mutate(date = as.Date(date)) %>%
  mutate_at(vars(town, county), str_to_title) %>%
  mutate(age_grp = cut(age, breaks = c(0, seq(10, 85, by = 5), 100), right = FALSE) %>%
           fct_relabel(age_brks))

cdc_wts <- read_csv("../raw_data/cdc_us_standard_pop_weights.csv") %>%
  mutate(age_grp = as_factor(age_grp) %>%
           fct_relabel(str_replace, "plus", "99"))
```

```{r}
opioid <- bind_rows(
  drugs_indiv %>% mutate(name = "Connecticut", level = "state"),
  drugs_indiv %>%
    left_join(ct5_clusters, by = "town") %>%
    rename(name = cluster) %>%
    mutate(level = "regions")
) %>%
  mutate(is_opioid = ifelse(is_opioid, "opioids_only", "not_opioid"))

daily_ts <- bind_rows(
  opioid %>% mutate(is_opioid = "all_drugs"),
  opioid %>% filter(is_opioid == "opioids_only")
) %>%
  count(level, date, name, is_opioid, age_grp) %>%
  as_tsibble(index = date, key = c(level, name, is_opioid, age_grp)) %>%
  mutate_if(is.factor, as.character) %>%
  fill_gaps(n = 0, .full = TRUE) %>%
  left_join(pop_age, by = c("level", "name", "age_grp" = "age_brk")) %>% 
  left_join(cdc_wts, by = "age_grp") # join pop. weights

drugs_ts <- daily_ts %>%
  index_by(month = yearmonth(date)) %>% 
  group_by(level, name, is_opioid, age_grp, pop, cdc_age_wt) %>%
  summarise(n = sum(n)) %>%
  mutate(rate = n / pop * 1e6) %>%
  group_by(level, name, is_opioid) %>%
  summarise(monthly_n = sum(n),
            monthly_age_adj = sum(rate * cdc_age_wt)) %>% 
  group_by_key() %>%
  mutate(rolling_mean_6mo = slider::slide_dbl(monthly_age_adj, mean, .before = 6, .complete = TRUE)) %>%
  # mutate(rolling_mean_6mo = ifelse(is.na(rolling_mean_6mo_0), monthly_age_adj, rolling_mean_6mo_0)) %>% 
  ungroup()
```

```{r}
yrs <- 2015:2020

annual_15_20 <- daily_ts %>%
  index_by(year = year(date)) %>% 
  filter(between(year, !!!range(yrs))) %>%
  as_tibble() %>%
  group_by(level, name, is_opioid, age_grp, pop, cdc_age_wt) %>%
  summarise(n = sum(n)) %>%
  mutate(ann_rate = n / pop * 1e5 / length(yrs)) %>%
  group_by(level, name, is_opioid) %>%
  summarise(total_n = sum(n),
            avg_annual_age_adj = sum(ann_rate * cdc_age_wt)) %>% 
  ungroup() %>%
  mutate(level = as_factor(level) %>% fct_relevel("state")) %>%
  arrange(level)
```

write files
```{r}
drugs_ts %>%
  mutate_at(vars(level, is_opioid), as_factor) %>%
  mutate(level = fct_relevel(level, "state")) %>%
  arrange(level, name) %>%
  saveRDS("../output_data/drug_overdose_5ct_monthly_ts_1mil.rds")

write_csv(annual_15_20, "../output_data/annual_avg_drug_deaths_100k_2015_2020.csv")
```


## Camille

```{r}
font_add_google("Roboto Condensed", "roboto")
showtext_auto()
tol12 <- ggthemes::ptol_pal()(12)
source("../_utils/geographies.R")

region_df <- cwi::regions %>%
  enframe(name = "region", value = "town") %>%
  unnest()
```

```{r}
age_brks <- function(x) {
  splits <- x %>%
    str_remove_all("[\\[\\]\\(\\)]") %>%
    str_split(",")
  splits %>%
    map_chr(function(xs) {
      x1 <- xs[1]
      x2 <- as.numeric(xs[2]) - 1
      sprintf("ages%s_%s", x1, x2)
    })
}

# for replacing with wider breaks for rates by race
age_wide_brks <- list(
  ages00_14 = c("ages00_09", "ages10_14"),
  ages15_24 = c("ages15_19", "ages20_24"),
  ages25_34 = c("ages25_29", "ages30_34"),
  ages35_44 = c("ages35_39", "ages40_44"),
  ages45_54 = c("ages45_49", "ages50_54"),
  ages55_64 = c("ages55_59", "ages60_64"),
  ages65_74 = c("ages65_69", "ages70_74"),
  ages75_84 = c("ages75_79", "ages80_84")
)
```


Need to do:

* Counts of overdose deaths
* Population by age to match age groups in CDC weights
* Rates
* Age-adjusted rates

```{r}
drugs_indiv <- read_csv("../output_data/overdoses_indiv_2012_2020.csv")
cdc_wts <- read_csv("../raw_data/cdc_us_standard_pop_weights.csv") %>%
  mutate(age_grp = as_factor(age_grp) %>%
           fct_relabel(str_replace, "plus", "over"))
```

```{r eval=FALSE, include=FALSE}
pop_read <- c(total = "", white = "H", black = "B", latino = "I") %>%
  map_chr(~paste0("B01001", .)) %>%
  map(~multi_geo_acs(table = ., year = 2018, regions = cwi::regions[main_regions])) %>%
  map(filter, !str_detect(level, "counties")) %>%
  map(label_acs)
saveRDS(pop_read, "../raw_data/pop_by_age_race_sex.rds")
```

```{r}
pop_read <- readRDS("../raw_data/pop_by_age_race_sex.rds")
# different age groupings for total pop & by race
pop_by_age <- pop_read[["total"]] %>%
  separate(label, into = c("total", "sex", "age"), sep = "!!") %>%
  filter(!is.na(age) | (is.na(age) & is.na(sex))) %>% 
  mutate(age = as_factor(age) %>%
           fct_relabel(str_replace_all, " (to|and) ", "_") %>%
           fct_relabel(str_remove, " years") %>%
           fct_relabel(~paste0("ages", .)) %>%
           fct_collapse(ages00_09 = c("agesUnder 5", "ages5_9"),
                        ages15_19 = c("ages15_17", "ages18_19"),
                        ages20_24 = c("ages20", "ages21", "ages22_24"),
                        ages60_64 = c("ages60_61", "ages62_64"),
                        ages65_69 = c("ages65_66", "ages67_69")) %>%
           fct_explicit_na(na_level = "total_pop") %>%
           fct_relevel("total_pop")) %>%
  mutate(level = fct_relabel(level, str_remove, "^\\d+_")) %>%
  group_by(level, name = NAME, age_grp = age) %>%
  summarise(pop = sum(estimate))
```

```{r}
pop_by_age_sex <- pop_read[["total"]] %>%
  separate(label, into = c("total", "sex", "age"), sep = "!!") %>%
  filter(!is.na(age)) %>%
  mutate(age = as_factor(age) %>%
           fct_relabel(str_replace_all, " (to|and) ", "_") %>%
           fct_relabel(str_remove, " years") %>%
           fct_relabel(~paste0("ages", .)) %>%
           fct_collapse(ages00_09 = c("agesUnder 5", "ages5_9"),
                        ages15_19 = c("ages15_17", "ages18_19"),
                        ages20_24 = c("ages20", "ages21", "ages22_24"),
                        ages60_64 = c("ages60_61", "ages62_64"),
                        ages65_69 = c("ages65_66", "ages67_69"))) %>%
  mutate(level = fct_relabel(level, str_remove, "^\\d+_"),
         sex = as.factor(sex) %>% fct_relabel(str_to_lower)) %>%
  group_by(level, name = NAME, sex, age_grp = age) %>%
  summarise(pop = sum(estimate))
```

```{r}
pop_by_age_race <- pop_read[c("white", "black", "latino")] %>%
  bind_rows(.id = "race") %>%
  separate(label, into = c("total", "sex", "age"), sep = "!!") %>%
  filter(!is.na(age)) %>%
  mutate(age = as_factor(age) %>%
           fct_relabel(str_replace_all, " (to|and) ", "_") %>%
           fct_relabel(str_remove, " years") %>%
           fct_relabel(~paste0("ages", .)) %>%
           fct_collapse(ages00_14 = c("agesUnder 5", "ages5_9", "ages10_14"),
                        ages15_24 = c("ages15_17", "ages18_19", "ages20_24"),
                        ages25_34 = c("ages25_29", "ages30_34"))) %>%
  mutate(level = fct_relabel(level, str_remove, "^\\d+_"),
         sex = as.factor(sex) %>% fct_relabel(str_to_lower)) %>%
  group_by(level, name = NAME, sex, race, age_grp = age) %>%
  summarise(pop = sum(estimate))

cdc_wt_race <- cdc_wts %>%
  mutate(age_grp = fct_collapse(age_grp, !!!age_wide_brks)) %>%
  group_by(age_grp) %>%
  summarise_all(sum)
```

Make a time series object with lots of groupings included (sex, age), then rollup later when I don't need disaggregations. Keeping a separate time series for counts by race, since ages are different.

*6/21*: adding in large towns with high counts. Filtering for towns with average 25 deaths per year--totally arbitrary, but gets 5 major towns to highlight.

```{r message=TRUE, warning=TRUE}
city_min_n <- 25
cities <- drugs_indiv %>%
  group_by(town, year) %>%
  summarise(n = n()) %>%
  summarise(avg_n = mean(n)) %>%
  filter(avg_n >= city_min_n)
drugs_ts <- bind_rows(
  drugs_indiv %>% 
    mutate(name = "Connecticut", level = "state"),
  drugs_indiv %>% 
    inner_join(region_df, by = "town") %>% 
    rename(name = region) %>%
    mutate(level = "regions"),
  drugs_indiv %>% 
    semi_join(cities, by = "town") %>%
    rename(name = town) %>%
    mutate(level = "towns")
) %>%
  mutate(level = as_factor(level)) %>%
  # mutate_at(vars(region, sex), as_factor) %>%
  count(date, level, name, sex, age_grp, is_opioid, is_fentanyl, is_heroin, is_cocaine, is_amphet, is_xylazine) %>%
  as_tsibble(index = date, key = c(level, name, sex, age_grp, is_opioid, is_fentanyl, is_heroin, is_cocaine, is_amphet, is_xylazine)) %>%
  fill_gaps(n = 0, .full = TRUE)

# need to redo age breaks for by race
drugs_race_ts <- bind_rows(
  drugs_indiv %>% mutate(name = "Connecticut", level = "state"),
  drugs_indiv %>% inner_join(region_df, by = "town") %>% mutate(level = "regions") %>% rename(name = region)
) %>%
  filter(race != "other_unknown") %>%
  mutate(age_grp = fct_collapse(age_grp, !!!age_wide_brks)) %>%
  count(date, level, name, sex, race, age_grp) %>%
  as_tsibble(index = date, key = c(level, name, sex, race, age_grp), .drop = T) %>%
  fill_gaps(n = 0, .full = TRUE)

out <- list()
```

```{r}
out$trend_all <- drugs_ts %>%
  mutate_if(is.factor, as.character) %>%
  index_by(month = yearmonth(date)) %>% 
  group_by(level, name, age_grp) %>%
  summarise(n = sum(n)) %>% 
  inner_join(pop_by_age, by = c("name", "age_grp", "level")) %>% 
  mutate(rate_100k = (n / pop) * 1e5) %>%
  left_join(cdc_wts, by = "age_grp") %>%
  summarise(monthly_n = sum(n),
            monthly_age_adj = sum(rate_100k * cdc_age_wt)) %>%
  group_by_key() %>%
  mutate(rolling_mean_6mo = slider::slide_dbl(monthly_age_adj, mean, .before = 6, .complete = TRUE)) %>%
  ungroup()

out$trend_sex <- drugs_ts %>%
  filter(sex != "unknown") %>%
  mutate_if(is.factor, as.character) %>%
  index_by(month = yearmonth(date)) %>% 
  group_by(level, name, sex, age_grp) %>%
  summarise(n = sum(n)) %>% 
  inner_join(pop_by_age_sex, by = c("name", "level", "sex", "age_grp")) %>% 
  mutate(rate_100k = (n / pop) * 1e5) %>%
  left_join(cdc_wts, by = "age_grp") %>%
  summarise(monthly_n = sum(n),
            monthly_age_adj = sum(rate_100k * cdc_age_wt)) %>%
  group_by_key() %>%
  mutate(rolling_mean_6mo = slider::slide_dbl(monthly_age_adj, mean, .before = 6, .complete = TRUE)) %>%
  ungroup()
```

```{r}
out$trend_race <- drugs_race_ts %>%
  mutate_if(is.factor, as.character) %>%
  index_by(month = yearmonth(date)) %>% 
  group_by_key() %>%
  summarise(n = sum(n)) %>% 
  inner_join(pop_by_age_race, by = c("level", "name", "sex", "race", "age_grp")) %>% 
  group_by(level, name, race, age_grp) %>%
  summarise_at(vars(n, pop), sum) %>%
  mutate(rate_100k = (n / pop) * 1e5) %>%
  left_join(cdc_wt_race, by = "age_grp") %>%
  summarise(monthly_n = sum(n),
            monthly_age_adj = sum(rate_100k * cdc_age_wt)) %>%
  group_by_key() %>%
  mutate(rolling_mean_6mo = slider::slide_dbl(monthly_age_adj, mean, .before = 6, .complete = TRUE)) %>%
  ungroup()
```


```{r}
out$trend_opioid <- drugs_ts %>%
  mutate_if(is.factor, as.character) %>%
  index_by(month = yearmonth(date)) %>% 
  group_by(level, name, is_opioid, age_grp) %>%
  summarise(n = sum(n)) %>% 
  inner_join(pop_by_age, by = c("level", "name", "age_grp")) %>% 
  mutate(rate_100k = (n / pop) * 1e5) %>%
  left_join(cdc_wts, by = "age_grp") %>%
  summarise(monthly_n = sum(n),
            monthly_age_adj = sum(rate_100k * cdc_age_wt)) %>%
  group_by_key() %>%
  mutate(rolling_mean_6mo = slider::slide_dbl(monthly_age_adj, mean, .before = 6, .complete = TRUE))

out$trend_fentanyl <- drugs_ts %>%
  mutate_if(is.factor, as.character) %>%
  index_by(month = yearmonth(date)) %>% 
  group_by(level, name, is_fentanyl, age_grp) %>%
  summarise(n = sum(n)) %>% 
  inner_join(pop_by_age, by = c("level", "name", "age_grp")) %>% 
  mutate(rate_100k = (n / pop) * 1e5) %>%
  left_join(cdc_wts, by = "age_grp") %>%
  summarise(monthly_n = sum(n),
            monthly_age_adj = sum(rate_100k * cdc_age_wt)) %>%
  group_by_key() %>%
  mutate(rolling_mean_6mo = slider::slide_dbl(monthly_age_adj, mean, .before = 6, .complete = TRUE))

out$trend_heroin <- drugs_ts %>%
  mutate_if(is.factor, as.character) %>%
  index_by(month = yearmonth(date)) %>% 
  group_by(level, name, is_heroin, age_grp) %>%
  summarise(n = sum(n)) %>% 
  inner_join(pop_by_age, by = c("level", "name", "age_grp")) %>% 
  mutate(rate_100k = (n / pop) * 1e5) %>%
  left_join(cdc_wts, by = "age_grp") %>%
  summarise(monthly_n = sum(n),
            monthly_age_adj = sum(rate_100k * cdc_age_wt)) %>%
  group_by_key() %>%
  mutate(rolling_mean_6mo = slider::slide_dbl(monthly_age_adj, mean, .before = 6, .complete = TRUE))
```

```{r}
out$trend_cocaine <- drugs_ts %>%
  mutate_if(is.factor, as.character) %>%
  index_by(month = yearmonth(date)) %>% 
  group_by(level, name, is_cocaine, age_grp) %>%
  summarise(n = sum(n)) %>% 
  inner_join(pop_by_age, by = c("level", "name", "age_grp")) %>% 
  mutate(rate_100k = (n / pop) * 1e5) %>%
  left_join(cdc_wts, by = "age_grp") %>%
  summarise(monthly_n = sum(n),
            monthly_age_adj = sum(rate_100k * cdc_age_wt)) %>%
  group_by_key() %>%
  mutate(rolling_mean_6mo = slider::slide_dbl(monthly_age_adj, mean, .before = 6, .complete = TRUE))

out$trend_amphet <- drugs_ts %>%
  mutate_if(is.factor, as.character) %>%
  index_by(month = yearmonth(date)) %>% 
  group_by(level, name, is_amphet, age_grp) %>%
  summarise(n = sum(n)) %>% 
  inner_join(pop_by_age, by = c("level", "name", "age_grp")) %>% 
  mutate(rate_100k = (n / pop) * 1e5) %>%
  left_join(cdc_wts, by = "age_grp") %>%
  summarise(monthly_n = sum(n),
            monthly_age_adj = sum(rate_100k * cdc_age_wt)) %>%
  group_by_key() %>%
  mutate(rolling_mean_6mo = slider::slide_dbl(monthly_age_adj, mean, .before = 6, .complete = TRUE))

out$trend_xyl <- drugs_ts %>%
  mutate_if(is.factor, as.character) %>%
  index_by(month = yearmonth(date)) %>% 
  group_by(level, name, is_xylazine, age_grp) %>%
  summarise(n = sum(n)) %>% 
  inner_join(pop_by_age, by = c("level", "name", "age_grp")) %>% 
  mutate(rate_100k = (n / pop) * 1e5) %>%
  left_join(cdc_wts, by = "age_grp") %>%
  summarise(monthly_n = sum(n),
            monthly_age_adj = sum(rate_100k * cdc_age_wt)) %>%
  group_by_key() %>%
  mutate(rolling_mean_6mo = slider::slide_dbl(monthly_age_adj, mean, .before = 6, .complete = TRUE))
```

## Towns
```{r}
yrs <- 2015:2020
min_pop <- 3000
ann_age_adj_15_20 <- bind_rows(
  drugs_indiv %>% mutate(name = "Connecticut", level = "state"),
  drugs_indiv %>% inner_join(region_df, by = "town") %>% rename(name = region) %>% mutate(level = "regions"),
  drugs_indiv %>% rename(name = town) %>% mutate(level = "towns")
) %>%
  ungroup() %>%
  filter(year %in% yrs) %>%
  mutate(is_opioid = if_else(is_opioid, "opiate_opioid", "no_opiate_opioid")) %>%
  bind_self(group = is_opioid, new_value = "any_substance", append = F) %>%
  count(level, name, is_opioid, age_grp) %>%
  complete(nesting(level, name), age_grp, is_opioid, fill = list(n = 0)) %>%
  inner_join(pop_by_age %>%
              ungroup() %>%
              select(-level), by = c("name", "age_grp")) %>%
  mutate(ann_rate100k = (n / pop) * 1e5 / length(yrs)) %>%
  left_join(cdc_wts, by = c("age_grp")) %>%
  group_by(level, name, is_opioid) %>%
  summarise(n = sum(n),
            pop = sum(pop),
            ann_crude_rate100k = (n / pop) * 1e5 / length(yrs),
            ann_age_adj_rate100k = sum(ann_rate100k * cdc_age_wt)) %>%
  mutate_at(vars(ends_with("rate100k")), round, digits = 1) %>%
  ungroup()

ann_age_adj_15_20 %>%
  filter(level == "towns", 
         is_opioid == "any_substance",
         n >= length(yrs), pop >= min_pop) %>%
  full_join(town_sf, by = "name") %>%
  mutate(rate_brk = classInt::classIntervals(ann_age_adj_rate100k, n = 5, style = "jenks")$brks %>%
           cut(ann_age_adj_rate100k, breaks = ., include.lowest = T)) %>%
  st_as_sf() %>%
  ggplot() +
  geom_sf(aes(fill = rate_brk), size = 0.2, color = "white") +
  coord_sf(ndiscr = 0) +
  theme_din() +
  scale_fill_brewer(palette = "Blues", na.value = "gray80") +
  labs(title = "Annualized age-adjusted overdose death rate, 2015-2019",
       fill = "Rate per 100k")
```

## Is cause of death changing?

```{r}
out$trend_opioid %>%
  filter(!str_detect(name, "Ring")) %>%
  select(-monthly_n) %>%
  gather(key, value, monthly_age_adj, rolling_mean_6mo) %>%
  ggplot(aes(x = month, y = value, color = is_opioid, alpha = key)) +
  geom_line() +
  scale_color_manual(values = tol12[c(1, 11)]) +
  scale_alpha_manual(values = c(monthly_age_adj = 0.3, rolling_mean_6mo = 0.9)) +
  theme_din() +
  facet_wrap(vars(name)) +
  labs(title = "Monthly age-adjusted rate of overdose deaths",
       subtitle = "By presence of opioid per 100k, with 6 month rolling average") +
  theme(legend.position = "bottom")
```




Have non-opioid deaths also gone up?

```{r}
drugs_indiv %>%
  count(year = year(date), is_opioid) %>%
  filter(year %in% range(year)) %>%
  spread(key = year, value = n, sep = "_") %>%
  mutate(percent_change = (year_2020 - year_2012) / year_2012)
```

Slightly. 17 percent increase in number of non-opioid deaths; 215 percent increase in number of opioid-related.

How has fentanyl factored in?

```{r}
out$trend_fentanyl %>%
  filter(!str_detect(name, "Ring")) %>%
  select(-monthly_n) %>%
  gather(key, value, monthly_age_adj, rolling_mean_6mo) %>%
  ggplot(aes(x = month, y = value, color = is_fentanyl, alpha = key)) +
  geom_line() +
  scale_color_manual(values = tol12[c(1, 11)]) +
  scale_alpha_manual(values = c(monthly_age_adj = 0.3, rolling_mean_6mo = 0.9)) +
  theme_din() +
  facet_wrap(vars(name)) +
  labs(title = "Monthly age-adjusted rate of overdose deaths",
       subtitle = "By presence of fentanyl per 100k, with 6 month rolling average") +
  theme(legend.position = "bottom")
```

```{r fig.width=7, fig.height=7}
month6_fentanyl_n <- drugs_ts %>%
  # filter(region == "Connecticut") %>%
  mutate_if(is.factor, as.character) %>%
  index_by(mo6 = floor_date(date, "6 months")) %>%
  group_by(level, name, is_fentanyl) %>%
  summarise(n = sum(n)) %>%
  as_tibble() %>%
  group_by(level, name, mo6)

month6_fentanyl_n %>%
  filter(name %in% str_subset(main_regions, "Ring", negate = T)) %>%
  mutate(share = n / sum(n)) %>%
  ggplot(aes(x = mo6, y = n, fill = is_fentanyl)) + 
  geom_col(alpha = 0.9) +
  geom_text(aes(label = scales::percent(share, accuracy = 1)), 
            data = . %>% filter(is_fentanyl),
            vjust = -0.1, family = "roboto", size = 2.8) +
  theme_din() +
  scale_fill_manual(values = tol12[c(3, 11)]) +
  scale_x_date(date_breaks = "1 year", date_labels = "%Y") +
  facet_wrap(vars(name), scales = "free_y", ncol = 2) +
  labs(title = "Count of deaths by presence of fentanyl",
       subtitle = "6-month totals, 2012-2020\nWith share of deaths that involve fentanyl",
       x = NULL) +
  theme(legend.position = "bottom")
```



```{r}
ann_age_adj_race_15_20 <- drugs_race_ts %>%
  filter(year(date) %in% yrs) %>%
  as_tibble() %>%
  group_by_at(vars(level:age_grp)) %>%
  summarise(n = sum(n)) %>%
  filter(sex != "unknown") %>%
  complete(nesting(level, name), sex, race, age_grp, fill = list(n = 0)) %>%
  inner_join(pop_by_age_race, by = c("level", "name", "sex", "race", "age_grp")) %>%
  group_by(level, name, race, age_grp) %>%
  summarise_at(vars(n, pop), sum) %>%
  mutate(rate_100k = (n / pop) * 1e5,
         ann_rate_100k = rate_100k / length(yrs)) %>%
  left_join(cdc_wt_race, by = "age_grp") %>%
  group_by(level, name, race) %>%
  summarise(n = sum(n),
            pop = sum(pop),
            ann_crude_rate100k = (n / pop) * 1e5 / length(yrs),
            ann_age_adj_rate100k = sum(ann_rate_100k * cdc_age_wt)) %>%
  ungroup() %>%
  mutate_at(vars(level, race), as.factor) %>%
  mutate_at(vars(ends_with("rate100k")), round, digits = 1) %>%
  mutate(level = fct_relevel(level, "state"),
         race = fct_relevel(race, "white", "black")) %>%
  arrange(level, name, race)
```

```{r}
ann_age_adj_race_15_20 %>%
  mutate(name = as.factor(name) %>% fct_reorder(ann_age_adj_rate100k, max)) %>%
  ggplot(aes(x = name, y = ann_age_adj_rate100k, color = race)) +
  geom_point(size = 2.5, alpha = 0.9) +
  coord_flip() +
  theme_din(xgrid = T) +
  labs(title = "Drug overdose rate by race",
       subtitle = "Annualized age-adjusted rate per 100k, 2015-2020") +
  scale_color_manual(values = tol12[c(1, 6, 3)])
```

## Output

```{r}
bind_rows(
  ann_age_adj_15_20 %>% mutate(type = "by_substance", race = "all"),
  ann_age_adj_race_15_20 %>% mutate(type = "by_race", is_opioid = "any_substance")
) %>%
  select(type, level, name, race, is_opioid, everything()) %>%
  mutate_at(vars(type, level, race, is_opioid), as_factor) %>%
  mutate(level = fct_relevel(level, "state"),
         race = fct_relevel(race, "all", "white")) %>%
  arrange(type, level, name, race, is_opioid) %>%
  filter(n >= length(yrs), pop >= min_pop) %>%
  rename_all(str_replace, "ann", "annualized") %>%
  split(.$type) %>%
  saveRDS("../output_data/age_adj_drug_deaths_2015_2020.rds")
```



```{r}
saveRDS(out, "../output_data/drug_overdose_age_adj_time_series_by_group_2012_2020.rds")
```

Adding in median age at time of death by race & by sex

```{r}
median_age <- drugs_indiv %>%
  filter(date > "2015-12-31") %>% 
  inner_join(region_df %>% filter(region %in% c("Fairfield County", "Greater New Haven", "Greater Hartford", "Lower Naugatuck Valley")),
            by = "town") %>%
  bind_rows(drugs_indiv %>% mutate(region = "Connecticut")) %>%
  bind_self(group = sex, new_value = "all") %>%
  bind_self(group = race, new_value = "all") %>%
  filter(sex != "unknown") %>%
  group_by(region, sex, race) %>%
  summarise(median_age = median(age, na.rm = T))

write_csv(median_age, "../output_data/median_age_drug_overdose_2016_2020.csv")
```

can't figure this out
```{r}
rates_od <- out$trend_all %>% 
  as_tibble() %>% 
	select(level, name, month, all = monthly_n) %>% 
	left_join(out$trend_opioid %>%  as_tibble() %>% 
							select(level, name, month, opioid = monthly_n),
						by = c("level", "name", "month")) %>% 
	left_join(out$trend_fentanyl %>%  as_tibble() %>% 
							select(level, name, month, fentanyl = monthly_n),
						by = c("level", "name", "month")) %>%  
	left_join(out$trend_heroin %>%  as_tibble() %>% 
							select(level, name, month, heroin = monthly_n),
						by = c("level", "name", "month")) %>% 
	left_join(out$trend_cocaine %>%  as_tibble() %>% 
							select(level, name, month, cocaine = monthly_n),
						by = c("level", "name", "month")) %>% 
	left_join(out$trend_amphet %>%  as_tibble() %>% 
							select(level, name, month, amphet = monthly_n),
						by = c("level", "name", "month")) %>% 
	ungroup() %>% 
	mutate(year = year(month)) %>% 
	select(-month) %>% 
	group_by(level, name, year) %>% 
	summarise_all(sum) %>% 
	ungroup() %>% 
	pivot_longer(cols = all:amphet, names_to = "substance", values_to = "count") %>% 
	group_by(level, name, year) %>% 
	camiller::calc_shares(group = substance, denom = "all", value = count) %>% 
	ungroup()

write_csv(rates_od, file = "../output_data/od_counts_and_rates_by_substance.csv")
```



